Using numerical modelling to investigate temperature and wave exposure as drivers

### 2020-01-22 thesis_chapter_1.Rmd
added to calculating drift section and updated all outputs

---
title: "Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model"
author: 
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  word_document:
    fig_caption: yes
    reference_docx: chapter_1_template.docx
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    theme:  simplex
bibliography: references.bib
---

@@

## Light

@@

## Depth

@@

## Nutrients

@@

## Temperature

@@

## Wave exposure

@@

# Particle dispersion modelling
## Particle dispersion models

@@

# Calculating characteristics of kelp drift
## Buoyancy of drifting objects in the oceans

@@

$V_W$ is the velocity of the object as a result of direct wave force, and is dependent on the size and shape of the object. In instances when wave-wave interactions result in an small, unbalanced force then $V_W$ is assumed to be zero. Also, objects may be able to absorb wave energy which can result in a non-zero $V_W$ value. Since surface-waves are wind driven, the $V_W$ is calculated by $$V_W = f(U_10)$$

>>>>

---
title: Quantifying passive kelp rafting phenomena and connectivity of nearshore communities
  model
  around the coast of South Africa:Development of a 3D spectral numerical coastal-flow
author:
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  word_document:
    fig_caption: yes
    reference_docx: markdown/thesis_template.docx
    toc: yes
  html_document:
    df_print: paged
    theme: simplex
    toc: yes
  pdf_document:
    toc: yes
bibliography: references.bib
---

@@

@@

@@

@@

@@

@@

# Drifting aspects of floating objects

The need to understand the effects of oceanographic conditions affecting trajectory of floating objects has largely been borne from the maritime industry. The applications in the maritime industry include locating lost cargo, locating naval and plane deris, search and rescue, and the hydrodynamic effects on naval architecture.

The trajectory of passivley drifting objects on the sea surface is influenced by multiple factors, such as water currents, atmospheric wind, wave motion, wave induced currents, gravitaional force and buoyancy force. To complicate matters, the previously mentioned factors do not act independently of one another but instead influence one another. Futhermore, the gravitational and buoyancy forces on the object are determined by the objects shape. Therefore, all these factors need to be taken into account when modelling trajectory of drifting objects. Given the local wind, surface current, and the shape and buoyancy of the object is known, it is possible to estimate trajectory by the equation $$V_{drift}=V_{curr} + V_{rel}$$,

where $V_{curr}$ represents the current velocity relative to the earth, and $V_{rel}$ represents the object drift velocity relative to the ambient water. Ocean currents are determined by two components: the surface current (including the effects of Ekman drift, baroclinic motion, tidal and inertial currents) and Stokes drift induced by waves. The assumption made is $V_{curr}$ influences all floating objects in the same manner, regardless of shape or size. Therefore, $V_{curr}$ is equated with the surface current obtained from a numerical ocean model which has been paraterised by wind velocity. The effects of $V_{rel}$ on a floating object is driven by wind and wave forces which is dependent on the shape and size of the floating object.

$V_W$ is the velocity of the object as a result of direct wave force, and is dependent on the size and shape of the object. In instances when wave-wave interactions result in an small, unbalanced force then $V_W$ is assumed to be zero. Also, objects may be able to absorb wave energy which can result in a non-zero $V_W$ value. Since surface-waves are wind driven, the $V_W$ is calculated by $$V_W = f(U_10)$$
# Particle dispersion modelling
## Particle dispersion models

@@

# Buoyancy of drifting objects in the oceans

@@

### 2020-01-22 thesis_chapter_2.Rmd
added to calculating drift section and updated all outputs

## Make first map of whole study area
map_A <- ggplot() +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")
ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
## Make 2nd map of just the peninsula
map_B <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

@@

## Create combined map
map_combined <- ggplot() +
  # First set the x and y axis limits so we know what the ranges are
  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
  # Then we place our facetsover one another using the coordinates we created
  annotation_custom(grob_A,
                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
  annotation_custom(grob_B,
                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
```

>>>>

## Make first map of whole study area
map_A <- ggplot() +
geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
            size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat), color = "red") +
  geom_text(data = study_sites_edit, aes(x = lon, y = lat, label = site), nudge_x = -0.1, nudge_y = -0.01) +
  annotate("text", label = "WARM\nTEMPERATE", x = 19.0, y = -34.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "COOL\nTEMPERATE", x =17.8, y = -33.5, size = 4.0, angle = 0, colour = "black") +
  annotate("text", label = "Cape\nPeninsula", x = 18.4, y = -34.5, size = 3.0, angle = 0, colour = "black") +
  scaleBar(lon = 17.8, lat = -34.45, distanceLon = 10, distanceLat = 10, distanceLegend = -10, dist.unit = "km",
  arrow.length = 10, arrow.distance = 15, arrow.North.size = 3) +
  coord_cartesian(xlim = c(17.75, 19.15), ylim = c(-32.50, -34.55)) +
  theme(axis.ticks = element_blank()) + # Remove tick marks
  xlab("Latitude") +
  ylab("Longitude")
ggsave(filename = "map_A.jpg", plot = last_plot(), device = "jpeg", path = "map/")
## Make 2nd map of just the peninsula
map_B <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = study_sites, aes(x = lon, y = lat, colour = "red"), show.legend = FALSE) +
  geom_text(data = study_sites, aes(x = lon, y = lat, label = site), nudge_x = 0.02, nudge_y = 0) +
  coord_cartesian(xlim = c(18.275, 18.525), ylim = c(-34.375, -34.100)) +
  theme(axis.ticks = element_blank(), # Remove tick marks
        axis.text = element_blank(), # Remove lat/ lon numbers
        axis.title = element_blank()) # Remove lat/ lon labels
#ggsave(filename = "map_B.jpg", plot = last_plot(), device = "jpeg", path = "map/")

@@

## Create combined map
map_combined <- ggplot() +
  # First set the x and y axis limits so we know what the ranges are
  coord_equal(xlim = c(1, 10), ylim = c(1, 10), ratio = 1, expand = TRUE, clip = "on") +
  # Then we place our facetsover one another using the coordinates we created
  annotation_custom(grob_A,
                    xmin = 1, xmax = 10, ymin = 1, ymax = 10) +
  annotation_custom(grob_B,
                    xmin = 4, xmax = 10, ymin = 4.5, ymax = 10)
#ggsave(filename = "map_combined.jpg", plot = last_plot(), device = "jpeg", path = "map/", width = 8, height = 8)
```
